### 2025-04-11 index.qmd
20250411 - add packages

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
  "nflreadr", "nflplotR", "nflfastR", "nfl4th", "gsisdecoder", "progressr",
  "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", "ggdag",
  "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", "pROC",
  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", "AICcmodavg",
  "rstan", "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", "viridis",
  "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", "stringi", #"webshot2", 
  "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin",
  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot" #"mgcv",
)

>>>>

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "dplyr", "ggplot2", "tidyverse", "downlit", "xml2",
  "nflreadr", "nflplotR", "nflfastR", "nfl4th", "gsisdecoder", "progressr",
  "DescTools", "pwr", "pwrss", "WebPower", "XICOR", "dagitty", "ggdag",
  "lubridate", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", "pROC",
  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", "AICcmodavg",
  "rstan", "brms", "bbmle", "fitdistrplus", "mclust", "magrittr", "viridis",
  "plotly", "quantmod", "fPortfolio", "xts", "zoo", "forecast", "stringi", #"webshot2", 
  "parallel", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin",
  "caret", "gpboost", "cmdstanr", "learnr", "remotes", "corrplot", "correlation", 
  "mgcv", "rms", "car", "performance", "lavaan", "mice", "miceadds"
)

### 2025-04-11 multiple-regression.qmd
20250411 - add packages

We can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), and added-variable plots (@fig-addedVariablePlots), to evaluate for potential nonlinearity.
For evaluating linearity, we would expect minimal bend/curvature in the lines.

@@

To determine if residuals are correlated given the nested structure of the data, we can examine the proportion of variance that is attributable to the particular player.
To do this, we can estimate the intraclass correlation coefficient (ICC) from a [mixed model](#sec-mixedModels).

@@

@@

pairwiseRegression_fit <- lavaan(
  pairwiseRegression_syntax,
  sample.mean = varMeans,
  sample.cov = varCovariances,
  sample.nobs = numObs
)

@@

@@

Now, let's specify the imputation method:

@@

```{r}
imp <- mice(
  as.data.frame(dataToImpute),
  method = meth,
  predictorMatrix = pred,
  m = numImputations,
  maxit = 5, # generally use 100 maximum iterations; this example uses 5 for speed
  seed = 52242)
```

@@

@@

Please note that the $p$-value for regression coefficients assumes that the observations are independent—in particular, that the residuals are not correlated.
However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
This non-independence violates the traditional assumptions of the significance of regression coefficients.
We could address this assumption by analyzing only one season from each player or by estimating the significance of the regression coefficients using cluster-robust standard errors.
For simplicity in the models above, we present results above from the whole dataframe.
In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence.
Below, we demonstrate how to account for non-independence of observations using cluster-robust standard errors.

>>>>

We can also evaluate residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), and added-variable plots (@fig-addedVariablePlots) from the `car` package [@Fox2019; @R-car], to evaluate for potential nonlinearity.
For evaluating linearity, we would expect minimal bend/curvature in the lines.

@@

To determine if residuals are correlated given the nested structure of the data, we can examine the proportion of variance that is attributable to the particular player.
To do this, we can estimate the intraclass correlation coefficient (ICC) from a [mixed model](#sec-mixedModels) using the `performance` package [@Ludecke2021; @R-performance].

@@

We can estimate a regression model that uses pairwise deletion using the `lavaan` package [@Rosseel2012; @R-lavaan].

@@

pairwiseRegression_fit <- lavaan::lavaan(
  pairwiseRegression_syntax,
  sample.mean = varMeans,
  sample.cov = varCovariances,
  sample.nobs = numObs
)

@@

We can multiply impute data using the `mice` package [@vanBuuren2011; @R-mice].

@@

Now, let's specify the imputation method—we use the two-level predictive mean matching (`2l.pmm`) method from the `miceadds` package [@R-miceadds] to account for the nonindependent data (owing to multiple seasons per player):

@@

```{r}
imp <- mice::mice(
  as.data.frame(dataToImpute),
  method = meth,
  predictorMatrix = pred,
  m = numImputations,
  maxit = 5, # generally use 100 maximum iterations; this example uses 5 for speed
  seed = 52242)
```

@@

We can estimate a regression model that uses full information maximum likelihood using the `lavaan` package [@Rosseel2012; @R-lavaan].

@@

Please note that the $p$-value for regression coefficients assumes that the observations are independent—in particular, that the residuals are not correlated.
However, the observations are not independent in the `player_stats_seasonal` dataframe used above, because the same player has multiple rows—one row corresponding to each season they played.
This non-independence violates the traditional assumptions of the significance of regression coefficients.
We could address this assumption by analyzing only one season from each player or by estimating the significance of the regression coefficients using cluster-robust standard errors.
For simplicity in the models above, we present results above from the whole dataframe.
In @sec-mixedModels, we discuss [mixed model](#sec-mixedModels) approaches that handle repeated measures and other data that violate assumptions of non-independence.
Below, we demonstrate how to account for non-independence of observations using cluster-robust standard errors with the `rms` package [@R-rms].
